Management induced changes of soil organic carbon on global croplands

### 2020-10-06 SOCbudgetPaper/SOCbudget.tex
Mayor update of everything

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Web Ecology (we)
% Wind Energy Science (wes)

@@

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

@@

\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}

@@

Possible titles:

\begin{itemize}
\item
  How management changed soil organic carbon on global croplands
\item
  Agricultural soil have lost 37 GtC topsoil carbon
\item
  Linking agricultural management data to soil modeleling -- a global approach
\end{itemize}

Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and biospheric pools multiple times (Bathjes). As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref. permafrost melting). The specific amount of carbon stored in the soil is however uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996). The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools. Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles in this regard.\\
Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today. Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC (Lal, 2006). Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX). While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especial agricultural activities.\\
Field-scale models (ref. Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm activities. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.\\
Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands. This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system. We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation, and tillage. We thereby provide the first global, spatially explicit SOC loss maps that consider detailed agricultural management.\\
This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study. We continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices. Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines. In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing the climate zone specfic stock changes with default stock change factors from the IPCC Tier 1 approach and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects. Finally, we will discuss our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.\\
\newpage

\hypertarget{method}{%
\section{Method}\label{method}}

We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
In the following chapter we outine the most important relationships and assumptions. Table \ref{append:subsection2mrfunctions} provides further information on corresponding code within the R packages.

@@

\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}

In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
\begin{equation}
SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
\label{eq:inoutflow}
\end{equation}
with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cell indices and \(t\) years. We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them. Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.

@@

We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \ref{tab:datasourceinputs}). Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see \ref{sec:livstmanure}); for natural vegetation litterfall \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.

Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citet{ipcc_2019_2019}). For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients. This implies that both the amount of carbon as well as its structural composition determine the effective inflow. Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.

@@

\begin{equation}
\begin{aligned}
& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
\label{eq:decayrates}
\end{aligned}
\end{equation}

For cropland we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions. Data sources as well as drivers considered for each land-use types are shown in table \ref{tab:datasourcedecay}. To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.

@@

                      & Soil disturbance & tillage & [sec:tillage] \\
 \bottomhline
 \end{tabular}
 \belowtable{}
 \label{tab:datasourcedecay}
 \end{table*}

\hypertarget{soc-transfer-between-land-use-types}{%
\subsubsection{SOC transfer between land-use types}\label{soc-transfer-between-land-use-types}}

@@

\begin{equation}
SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
\label{eq:ctransfer}
\end{equation}

with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land-use type \(lut\). Note that \(!lut\) denotes the sum over all other land-use types, which decreases in the specific time step \(t\). Data sources and methodology on land-use states and changes are described in \ref{sec:landuse}.

\hypertarget{total-soc-stocks-and-stock-changes}{%
\subsubsection{Total SOC stocks and stock changes}\label{total-soc-stocks-and-stock-changes}}

@@

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
\label{eq:SOCstate}
\end{equation}

@@

\begin{equation}
SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
\label{eq:steadystate2budget}
\end{equation}

@@

\begin{equation}
SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
\label{eq:totalstock}
\end{equation}

@@

\begin{equation}
\Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
\label{eq:totalstock}
\end{equation}

Note that the absolution changes \(\Delta SOC\) can be also interpreted as cummulative SOC emissions or as SOC depth of mankind; whereas relative changes \(F^{SCF}\) can be denoted as stock change factors as defined within the IPCC guidelines of 2006 \citep{ipcc_2006_2006}.

\hypertarget{initialisation-of-soc-pools}{%
\subsubsection{Initialisation of SOC pools}\label{initialisation-of-soc-pools}}

@@

Land-use patterns are based on the Land-Use Harmonization 2 \citep{Hurtt2020} data set, which we aggregate from quarter degree to half degree resolution. We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see \ref{append:Tableluh2fao2mag} for more details on the crop group mapping). Land-use transitions are calculated as net area differences of the land-use data on half-degree.

@@

Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns. We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}. Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.

@@

\begin{equation}
\begin{aligned}
AGR & = CP \cdot HI_{\textrm{slope}} + CA \cdot HI_{\textrm{intercept}}\qquad & \textrm{and} \\
BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
\label{eq:resbiomass}
\end{aligned}
\end{equation}

@@

We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \ref{append:Tablekcr2kres}). Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see \citep{weindl}), taking available AGR biomass as well as livestock productivity into account. We estimate, for low-income countries, a material-use share for straw residues of 5\% and a fuel-share of 10\% for all used residues groups in low-income countries. For high-income countries, no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income. The remaining AGR as well as all BGR are assumend to be recycled to the soil. We limit high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.

@@

To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \ref{append:Tablec2dm}).

@@

\textbf{Manure Excretion, Storage and Recycling}
Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modelling. Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets \citep{weindl}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep{ipcc_2019_2019}).
Depending on the feed system we assume manure to be handled in four different ways:
All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel. Manure fuel shares are estimated using IPCC default values \citep{ippc_2006_2006}.
Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing.
To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled. For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see \ref{append:TableclossAWMS}). We limit high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.

@@

To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage. This rule is based on the approach of \citep{porwollik_generating_2018}, but neglects the (less common) adoption of no-tillage farming on agricultural land.

@@

Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}). Here, stocks are calculated via stock change factors (F\^{}\{SCF\}) given by the IPCC for the topsoil (0-30 cm) and based on observational data. \(F^{SCF}\) are differentiated by different crop, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these. Moreover \(F^{SCF}\) vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}). The actual SOC stocks as thus calulated based on a given reference SOC stock by

\begin{equation}
SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
\label{eq:tier1}
\end{equation}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

We present simulation results of the SOC budget of the period from 1975 to 2010 mainly focusing on cropland areas. Detailed total global and regional results of the state of the worlds SOC in comparison to other estimates can be found in the supplement.

\hypertarget{soc-distribution-and-depletion}{%
\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}

Figure

\begin{figure}[H]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas. (b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades (despite ongoing cropland expansion). (c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics. Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas.}\label{fig:SOCmaps}
\end{figure}

\hypertarget{global-agricultural-carbon-flows-and-stocks}{%
\subsection{global agricultural carbon flows and stocks}\label{global-agricultural-carbon-flows-and-stocks}}

>>>>

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

@@

@@

\correspondence{Kristine\ Karstens\ (kristine.karstens@pik-potsdam.de)}

@@

Soil Organic Carbon (SOC), the amount of organic carbon stored through
the earth's soil, is the largest terrestrial carbon pool, exceeding the
carbon in the atmospheric and vegetation pools multiple times (Batjes).
As such, even small changes in drivers of SOC may thus lead to
substantial shifts in earth carbon cycle and influence the amount of CO2
in the atmosphere (ref. permafrost melting). The specific amount of
carbon stored in the soil is, however, uncertain, with estimates ranging
from 1500 to 2400 GtC for the first meter of the soil profile (Batjes,
1996; more). The quality of SOC maps has markedly improved in recent
decades, along with better understanding of the factors driving the
global magnitude, distribution, and dynamics of SOC pools. Natural
properties like climatic, biophysical, and landscape characteristics
clearly play the most important roles to determine SOC variations over
space and time. Human intervention, including land cover change and land
management, has however added a further driver to SOC change, which
alters terrestrial carbon pools in much shorter time scales and is
likely one of the most dominant driver of SOC changes on managed land
today (cite). Recent studies identify the anthropogenic SOC debt for the
first meter of the soil profile at around 116 GtC (Sanderman et al.),
compared to previous estimates between 60-130 GtC (Lal, 2006). Other
studies have focused more closely on spatially disaggregation of SOC
changes via advanced digital soil mapping techniques (S-World;
Stoorvogel 2, 2017) or better representation of biogeochemical processes
within SOC dynamics (Hararuk, 20XX). While providing a rough estimate of
the order of magnitude of change, these studies lack a detailed
consideration of land management, especially of agricultural activities.
Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to
capture these land management impacts by using detailed information on
crop yield levels, fertilizer inputs and various other on-farm measures.
However, due to the lack of comprehensive global management data as
input to these models, scaling up to the global extent remains a complex
challenge. Our study combines an spatially explicit estimate of
agricultural management data on the global level with a 3-pool SOC model
parametrized for global croplands. This allows us to estimate SOC stock
change factors, as well as organic carbon flow dynamics within the
agricultural system. We thereby consider change in SOC caused by
historical land cover change as well as of different agricultural
management practices, including residue recycling, manure amendments,
irrigation, and tillage. We provide the first global, spatially explicit
SOC loss maps that considers detailed agricultural management. This
paper will first introduce in the methods section the basic concept of
SOC dynamics as applied in this study. We continue with a detailed
description of the global gridded management data used here, including
crop production levels, residue recycling rates, manure amendments, and
the adoption of irrigation and tillage practices. Lastly, we shortly
refer to the concept of stock change factors as outlined in the Tier 1
approach of the IPCC guidelines. In our results section we focus on the
SOC dynamics of global croplands by (1) analysing the spatially explicit
distribution and depletion of SOC from 1975 to 2010, (2) comparing
impacts of different management effects on SOC emissions and (3)
quantify global agricultural carbon flows and stocks to compare the
importance of various management aspects. Finally, we discuss and
compare our findings -- including their implications for SOC model
development -- and conclude with an outlook on the ability of SOC
management to mitigate climate change and contribute to negative
greenhouse gas emissions. \newpage

\section{Method}

We compile calculations as open-source R packages available at
github.com/pik-piam/mrcommons (management related functions) and
github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which
are both based on the MADRaT package (package citation?), a framework
which aims to improve reproducibility and transparency in data
processing. In the following chapter we outine the most important
relationships and assumptions. Table @ref(append:subsection2mrfunctions)
provides further information on corresponding code within the R
packages.

@@

\subsubsection{Steady-state SOC stocks and decay rates}

In a simple first order kinetic approach the steady-state soil organic
carbon stocks \(SOC^{eq}\) are given by \begin{equation}
SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
(\#eq:inoutflow)
\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil,
\(k\) denoting the soil organic carbon decay rate; as well as \(i\)
representing grid cell indices and \(t\) years. We use for our
calculations the Tier 2 steady-state method, which assumes three soil
carbon sub-pools (active, slow and passive) and interactions between
them. Annual carbon inflow to each sub-pool and annual decay rates of
each sub-pool are the key components to determining steady-state SOC
stocks.

@@

We account for different carbon input sources depending on the two
land-use types we distinguish: croplands and natural vegetated land as
representative for all other land use (see table
@ref(tab:datasourceinputs)). Carbon sources for cropland are recycled
crop residues, below ground biomass of crops (for both see
@ref(sec:residues)) and recycled manure (see @ref(sec:livstmanure)); for
natural vegetation litterfall \citep{schaphoff_lpjml4_2018} is the only
source of carbon inflow to the soil.

Following the IPCC carbon accounting methodology, carbon inputs are
disaggregated into metabolic and structural components depending on
their lignin and nitrogen content (see \citet{ipcc_2019_2019}). For each
component the sum of all carbon input sources is allocated to the
respective SOC sub-pools via transfer coefficients. This implies that
both the amount of carbon as well as its structural composition
determine the effective inflow. Data sources for all considered carbon
inputs as well as for lignin and nitrogen content can be found in table
@ref(tab:datasourceinputs).

@@

\begin{equation}
\begin{aligned}
& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
(\#eq:decayrates)
\end{aligned}
\end{equation}

For cropland we distinguish the effect of different tillage (see
@ref(sec:tillage)) and irrigation (see @ref(sec:irrigation)) practices
on decay rates, whereas for natural vegetation, we assume rainfed and
non-tilled conditions. Data sources as well as drivers considered for
each land-use types are shown in table @ref(tab:datasourcedecay). To
account for variations of decay rates within each grid cell due to
different tillage and irrigation regimes, average rates based on area
shares are calculated.

@@

                      & Soil disturbance & tillage & [sec:tillage] \\
 \bottomhline
 \end{tabular}
 \belowtable{}
 (\#tab:datasourcedecay)
 \end{table*}

\subsubsection{SOC transfer between land-use types}

@@

\begin{equation}
SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
(\#eq:ctransfer)
\end{equation}

with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
expansion for a given land-use type \(lut\). Note that \(!lut\) denotes
the sum over all other land-use types, which decreases in the specific
time step \(t\). Data sources and methodology on land-use states and
changes are described in @ref(sec:landuse).

\subsubsection{Total SOC stocks and stock changes}

@@

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
(\#eq:SOCstate)
\end{equation}

@@

\begin{equation}
SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
(\#eq:steadystate2budget)
\end{equation}

@@

\begin{equation}
SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
(\#eq:totalstock)
\end{equation}

@@

\begin{equation}
\Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
(\#eq:totalstock)
\end{equation}

Note that the absolution changes \(\Delta SOC\) can be also interpreted
as the SOC gap due to human cropping activits; whereas relative changes
\(F^{SCF}\) can be denoted as stock change factors as defined within the
IPCC guidelines of 2006 \citep{ipcc_2006_2006}. Note that the SOC gap is
to the negated cummulative SOC emission also refered to as SOC debt
(cite Sanderman) of mankind.

\subsubsection{Initialisation of SOC pools}

@@

Land-use patterns are based on the Land-Use Harmonization 2
\citep{Hurtt2020} data set, which we aggregate from quarter degree to
half degree resolution. We disaggregate the five different cropland
subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17
crop groups, applying the relative shares for each grid cell based on
the country and year specific area harvested shares of FAOSTAT data
\citep{faostat_faostat_2016} (see @ref(append:Tableluh2fao2mag) for more
details on the crop group mapping). Land-use transitions are calculated
as net area differences of the land-use data on half-degree.

@@

Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as
well as half-degree cropland patterns (see @ref(sec:landuse)) we compile
crop group specific half-degree production patterns. We calibrate
cellular yields with a country-level calibration factor for each crop
group to meet historical FAOSTAT production
\citep{faostat_faostat_2016}. Note that by using physical cropland areas
we account for multiple crop harvest events as well as for fallows.

@@

\begin{equation}
\begin{aligned}
AGR & = CP \cdot HI_{\textrm{slope}} + CA \cdot HI_{\textrm{intercept}}\qquad & \textrm{and} \\
BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
(\#eq:resbiomass)
\end{aligned}
\end{equation}

@@

We compile out of our 17 crop groups, three residue groups (straw,
high-lignin and low-lignin residues) with additional demand for other
purposes and one residue with no double use (see
@ref(append:Tablekcr2kres)). Residue feed demand for five different
livestock groups is based on country- and residue-group-specific feed
baskets (see \citep{weindl}), taking available AGR biomass as well as
livestock productivity into account. We estimate, for low-income
countries, a material-use share for straw residues of 5\% and a
fuel-share of 10\% for all used residues groups in low-income countries.
For high-income countries, no withdrawal for material or fuel use is
assumend, and middle-income countries use shares are linearly
interpolated based on per-capita income. The remaining AGR as well as
all BGR are assumend to be recycled to the soil. We limit high recycling
shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.

@@

To transform dry matter estimates into carbon, we compiled crop-group
and plant part specific carbon to dry matter (c:dm) ratios (see
@ref(append:Tablec2dm)).

@@

\textbf{Manure Excretion, Storage and Recycling} Manure production and
management is based on a revised methodology of
\citep{bodirsky_n2o_2012} and is presented here due to its central role
in soil carbon modelling. Based on the gridded livestock distribution we
calculate excretions by estimating the nitrogen balance of the livestock
system on the basis of comprehensive livestock feed baskets
\citep{weindl}, assuming that all nitrogen in protein feed intake, minus
the nitrogen in the slaughter mass, is excreted. Carbon in excreted
manure is estimated by applying fixed C:N ratios (given by
\citep{ipcc_2019_2019}). Depending on the feed system we assume manure
to be handled in four different ways: All manure orginated from pasture
feed intake is excreted directly to pastures and rangelands (pasture
grazing), deducting manure collected as fuel. Manure fuel shares are
estimated using IPCC default values \citep{ippc_2006_2006}. Whereas for
low-income countries, we adopt a share of 25\% of crop residues in feed
intake directly consumend and excreted on crop fields (stubble grazing),
we do not consider any stubble grazing in high-income countries;
middle-income countries see linearly interpolated shares depending on
their per-capita income. For all other feed items we assume the manure
to be stored in animal waste management systems associated with
livestock housing. To estimate the carbon actually recycled to the soil,
we account for carbon losses during storage and recycling shares in
different animal waste management and grazing systems. Whereas we assume
no losses for pasture and stubble grazing, we consider that the manure
collected as fuel is not recycled. For manure stored in different animal
waste management systems we compiled carbon loss rates partly depending
on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}
(see @ref(append:TableclossAWMS)). We limit high application shares at
\(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.

@@

To account for the distribution of tillage to the three different
tillage types specfied by the IPCC - full tillage, reduced tillage and
no tillage -, we assume that all natural land and pastures are not
tilled, whereas as default annual crops are under full and perennials
under reduced tillage. Furthermore we assume no tillage in cropland
cells specified as no tillage cell based on the historic global gridded
tillage data set from Prowollik \citep{porwollik_generating_2018}.
Porwollik data set is extended by to the period of 1974-2010 by
combining country-level data on conservation agriculture (ca) area
values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the
methodolgy of Porwollik to identify potential no tillage cells.

@@

Additionally to the tier 2 approach of the refinement of IPCC guidelines
vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management
data coming with it, SOC changes can be estimated using the IPCC tier 1
approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006},
\citet{ipcc_2019_2019}). Here, stocks are calculated via stock change
factors (F\^{}\{SCF\}) given by the IPCC for the topsoil (0-30 cm) and
based on observational data. \(F^{SCF}\) are differentiated by different
crop, management and input systems (here summarized under \(m\))
reflecting different dynamics under changed in- and outflows without
explicitly tracking these. Moreover \(F^{SCF}\) vary for different
climatic zones (\(c\)) specified by the IPCC (see
@ref(append:climatemap)). The actual SOC stocks as thus calulated based
on a given reference SOC stock by

\begin{equation}
SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
(\#eq:tier1)
\end{equation}

@@

\section{Results}

We present simulation results of our SOC budget focusing on cropland
areas for the year 2010 as well as global trends of SOC stock changes
for the period from 1975 to 2010.

\subsection{SOC distribution and depletion}

\begin{figure}[H]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas. (b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades. (c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics. Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas.}\label{fig:SOCmaps}
\end{figure}

In fig.~1(a) we provide the first world map of SOC on croplands
considering real world management data on the global scale. Our
spatially explicit results moreover show hot spots of SOC losses as well
as gains in two different ways: 1. Absolute SOC changes (see fig.~1(c))
indicate areas with high importance for the global SOC emissions. The
might be driven by huge relative losses or a high natural stock, from
which even small deviations could lead to substantial losses. 2. To
attribute SOC losses to insufficient agricultural management relative
SOC changes (\(F^{SCF}\), see fig.~1(d)) are a helpful tool. They
indicate areas with huge difference in carbon inflows or SOC decay
compared to natural vegetation, that might be overcome due to improved
agricultural pratices.

\subsection{Agricultural management effects on SOC emissions and
cycling}
